---
title: "Programming Assignment #4 (PA4)"
subtitle: "Data Science for Linguists"
author: "Quartz Colvin"
institute: "Rutgers University"
date: "2016/12/12 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: ["rutgers", "rutgers-fonts"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
# Setup

- First, we need to load the libraries

```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: true

library("tidyverse")
library("ds4ling")
library("untidydata")
library("plot3D")

options(htmltools.dir.version = FALSE)

```


---
# The data

## Load dataset 

**(4.) Load the dataset from inside your index.Rmd file**

```{r}
#| label: load_dataset
#| message: false
#| warning: false
#| echo: true

fricatives <- read_csv("./data_raw/fricatives/fricatives.csv")

```

---

# The data 

## "Codebook" (reference information)

- obs: (character), observation number
- s_cog: (numeric), center of gravity measurement in Hz for [s].
- s_skewness: (numeric), skewness measurement for [s].
- sh_cog: (numeric), center of gravity measurement in Hz for [ʃ].
- sh_skewness: (numeric), skewness measurement for [ʃ].

---

# The data 

## Preview the data

```{r}
#| label: view_dataset

head(fricatives)

```

---
# The data

## Tidy this data

**(5.) Tidy the data**

```{r}
#| label: tidy_fricatives
#| message: true
#| warning: false
#| echo: true

tidy_fricatives <- fricatives |>
  pivot_longer(
    c(s_cog, sh_cog, s_skewness, sh_skewness),
    names_to = c("phoneme", "values"),
    names_sep = "_") |>
  pivot_wider(
    names_from = values,
    values_from = value
  )

```

---

# The data 

## View tidy data

- Did this work? Let's see:

```{r}
#| label: glimpse_tidy_fricatives
#| message: true
#| warning: false
#| echo: true

head(tidy_fricatives)

```

- Thank god

---

# The data

## Save data

**(5.) Save the newly tidied data as a csv file in "data_tidy"**

```{r}
#| label: fricatives_tidy_csv

write_csv(x = fricatives, file = "./data_tidy/tidy_fricatives.csv")

```


---

# Statistics

## Descriptive statistics

**(6.) Provide a table of descriptive statistics**

- Merlin brought up the idea of using a function for this and I liked that idea.

- So! Here is the function I'm using:

```{r}
#| label: desriptive_stats_fx 
#| message: true
#| warning: false
#| echo: true

stats_fx <- function(x) {
  cat("Mean:", mean(x), "\n")
  cat("Std:", sd(x), "\n")
  cat("Min:", min(x), "\n")
  cat("Max:", max(x), "\n")
  cat("Median:", median(x), "\n")
  cat("Variance:", var(x), "\n")
}

```

---
### Descriptive statistics

```{r}
#| label: descriptive_stats_tables
#| message: true
#| warning: false
#| echo: true

knitr::kable(stats_fx(tidy_fricatives$cog),
             caption = "Descriptive Statistics: Center of Gravity.")

knitr::kable(stats_fx(tidy_fricatives$skewness),
             caption = "Descriptive statistics: Skewness.")
```


---
# Statistics 

## Boxplots

**(7.) Make a boxplot of center of gravity as a function of phoneme.**

- The boxplot is on the next slide!

---
### Boxplot (cog and phoneme)

```{r}
#| label: boxplot_cog_phoneme
#| echo: false
#| warning: false

tidy_fricatives |> 
  ggplot() + 
  aes(x = phoneme, y = cog) + 
  geom_boxplot() +
  labs(x = 'Phoneme', y = 'Center of Gravity') +
  labs(title = "Center of Gravity by Phoneme.")
```


---
# Statistics

## Plot skewness

**(7.) In another slide, plot skewness as a function of phoneme.**

- Boxplot is on the next slide.

---

### Boxplot (skewness and phoneme)

```{r}
#| label: boxplot_skewness_phoneme
#| echo: false
#| warning: false

tidy_fricatives |> 
  ggplot() + 
  aes(x = phoneme, y = skewness) + 
  geom_boxplot() +
  labs(x = 'Phoneme', y = 'Skewness') +
  labs(title = "Skewness by Phoneme.")

```

---
# Statistics

## Stat summary

**(7.) Use a statistical transformation (i.e., not a boxplot, but rather stat_summary()).**

- This plot is on the next slide.

---
### Stats summary

```{r}
#| label: fricatives_transformation
#| message: false
#| warning: false
#| echo: false

tidy_fricatives |> 
  ggplot() + 
  aes(x = phoneme, y = skewness) +
  stat_summary(
    fun.data = mean_sdl,
    geom = 'pointrange'
  ) +
  labs(x = "Phoneme", y = "Skewness") +
  labs(title = "Skewness by Phoneme ('stat_summary' version)")

```

---

# Model

## Fit a model 

**(8.) Fit a model that examines center of gravity as a function of skewness for the [s] segments (hint: you will have to transform the data).**

```{r}
#| label: mod_cog_skew

s_data <- tidy_fricatives |>
  filter(phoneme == "s") |>
  select(phoneme, cog, skewness)

s_mod <- lm(formula = cog ~ skewness, data = s_data)

```

---

# Model

## Model summary 

**(8.) Make a table of the model summary.**

```{r}
#| label: s_mod_summary
#| message: false
#| warning: false
#| echo: true

summary(s_mod)

```

---

# Model

## Make scatterplot

**(9.) Make a scatter plot that illustrates the relationship in the previous slides**

- See next slide.

---
### Scatterplot

```{r}
#| label: scatterplot_s_mod
#| message: false
#| warning: false
#| echo: true

s_data |> 
  ggplot() + 
  geom_point(mapping = aes(x = cog, y = skewness)) +
  geom_smooth(method='lm', aes(x = cog, y = skewness), se = FALSE) +
  labs(x = 'Center of gravity', y = 'Skewness') +
  labs(title = "Correlation between center of gravity and skewness for 's' phoneme.")

```


---

# Model

## Model diagnostics 

**(10.) Check model diagnostics (make plots).**

- See next slide

---
### Diagnostic plots

```{r}
#| label: diagnostics
#| message: false
#| warning: false
#| echo: true
#| fig-dpi: 300
#| fig-asp: 0.56
#| out-width: "100%"

ds4ling::diagnosis(s_mod)

```


---

### Results 

**(11.) Write up the results (as if it were for a journal article).**

- This model is fit **fairly well**.

- **Plot 1:** The Residuals plot shows that the points are somewhat narrowly spread out around 0. This could be a bit better, but for now it is ***pretty good***.

- **Plot 2:** The Density histogram is almost bell-shaped. This diagnostic *could* be better, but for now it is ***pretty good***. 

- **Plot 3:** The Simple Quartiles plot shows that the tail ends of the data don't deviate much from the baseline. Again, this diagnostic ***could be better***, but it isn't terrible. 

- So, overall this model is fit fairly well (especially with the consideration that no model is flawless) even though the plots could look a bit better.




---

# Assumptions 

## Load dataset

**(12.) In a new slide, load the assumptions.csv dataset.**

```{r}
#| label: load_assumptions
#| message: false
#| warning: false
#| echo: true

assumptions <- read_csv("./data_raw/fricatives/assumptions.csv")

```
---
# Assumptions dataset

## View the dataset

```{r}
#| label: view_assumptions
#| message: false
#| warning: false
#| echo: true

head(assumptions)

```

---

# Assumptions dataset 

## Scatterplot

**(12.) Make a scatterplot.**

- Oh!! A ***FRIEND***-shaped plot! 

---
### Our dinosaur friend

```{r}
#| label: scatterplot_assumptions
#| message: false
#| warning: false
#| echo: true

assumptions |>
  ggplot() + 
  aes(x = x, y = y) + 
  geom_point()

```


---
# Assumptions dataset 

## Fit model to this data?

**(12.) Explain in a few sentences why it would not be appropriate to fit a model to this data.**

- Let's run the diagnoses to see exactly why it wouldn't be reasonable.

- These plots are on the next slide.

---
### Fit a model?

```{r}
#| label: assumptions_diagnostics
#| message: false
#| warning: false
#| echo: true
#| fig-dpi: 300
#| fig-asp: 0.56
#| out-width: "100%"

ass_mod <- lm(y ~ x, data = assumptions)

ds4ling::diagnosis(ass_mod)

```

---

### (12.) Fit a model? ***No.*** 

Why is it not appropriate to fit a model to this data?

- **Plot 1:** In the Residuals plot, the points are spread out ***very randomly***. 

- **Plot 2:** The Density histogram is ***not bell-shaped*** at all. 

- **Plot 3:** The Simple Quartiles plot shows that the ends of the datapoints ***greatly deviate*** from the line. 

- All of these diagnoses indicate that **a fitted model isn't appropriate** for this dataset.

---

# "Thank you!" dino friend.

```{r}
#| label: dino_friend
#| message: false
#| warning: false
#| echo: false

assumptions |>
  ggplot() + 
  aes(x = x, y = y, color = "red") + 
  geom_point()

```

